initial_analysis_notebook

### 2021-04-27 exploratory/notebook/initial_analysis/initial_analysis_notebook.Rmd
setup lefse plot for other comparisons

lefse_df <- data.table::fread(paste0('../../../data/process/severe_disease_Genus.0.03.lefse_summary'),
		fill = T) %>% 
	filter(!is.na(LDA)) %>% 
	top_n(20, LDA) %>% 
	mutate(order = LDA + 10 * as.logical(Class),
		tax_otu_label = OTU)

lefse_plot <- lefse_df %>% 
	ggplot(aes(y = LDA, fill = Class, 
			x = reorder(tax_otu_label, order))) +
		geom_col() + 
		theme_bw() + 
		coord_flip() + 
		guides(fill = 'none') + 
		scale_fill_manual(values = c('#8de4d3','#f75b5f'),
			limits = c(TRUE, FALSE),
			labels = c('Severe', 'Non-severe')) + 
		labs(x = NULL, y = 'LDA Score (log 10)')

abundance_plot <- metadata %>% 
	filter(cdiff_strain == 431,
		day == 0) %>% 
	left_join(shared, by = c('group' = 'Group')) %>% 
	gather(OTU, abundance, matches('Otu\\d*')) %>% 
	left_join(taxonomy, by = 'OTU') %>% 
	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
	group_by(mouse_id, group, Genus, order, early_euth) %>% 
	summarise(abundance = sum(abundance)) %>% 
	mutate(tax_otu_label = Genus) %>% 
	mutate(abundance = abundance / 21.07,
		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
  highlight_key(~mouse_id) %>% 
	ggplot(aes(x = reorder(tax_otu_label, order), 
			y = abundance, color = as.factor(early_euth), shape = mouse_id)) + 
		stat_summary(fun.data = 'median_hilow', aes(group = early_euth),
			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
		theme_bw() + labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
			labels=c("0","0.1","1","10","100")) +
		scale_color_manual(values = c('#8de4d3','#f75b5f'),
						   limits = c(TRUE, FALSE),
						   labels = c('Severe', 'Non-severe')) + 
    scale_shape_manual(values = rep(16, 54)) + 
		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
		theme(panel.grid.major.y = element_blank(), 
			panel.grid.minor = element_blank(),
			legend.position = 'right') + 
		coord_flip() + 
    guides(shape = 'none')

cowplot::plot_grid(lefse_plot, 
                   abundance_plot + 
                     geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.25) + 
                     theme(axis.ticks.y = element_blank(),
                           axis.text.y = element_blank()), 
                   nrow = 1)
#library(plotly)
#gg <- ggplotly(abundance_plot + geom_jitter(position = position_jitterdodge(dodge.width = 0.5)), tooltip = c('mouse_id'))
#highlight(hide_legend(gg), on = "plotly_click", off = "plotly_doubleclick", 
#           color = "black", selectize = T, opacityDim = .2,
#           selected = attrs_selected(showlegend = F, marker = list(symbol = 'X')))

```
## Severe disease
```{r}
plot_lefse_data('severe_disease')
```
```

>>>>

plot_lefse_data <- function(file_name){
    lefse_design <- data.table::fread(paste0('../../../data/process/', file_name, '_Genus.design'))
    colnames(lefse_design) <- c('Group', 'class')

    lefse_df <- data.table::fread(paste0('../../../data/process/', file_name, '_Genus.0.03.lefse_summary'),
  		fill = T) %>% 
  	filter(!is.na(LDA)) %>% 
  	top_n(20, LDA) %>% 
  	mutate(order = LDA + 10 * as.logical(Class),
  		tax_otu_label = OTU)

  lefse_plot <- lefse_df %>% 
  	ggplot(aes(y = LDA, fill = Class, 
  			x = reorder(tax_otu_label, order))) +
  		geom_col() + 
  		theme_bw() + 
  		coord_flip() + 
  		guides(fill = 'none') + 
  		scale_fill_manual(values = c('#8de4d3','#f75b5f')) + 
  		labs(x = NULL, y = 'LDA Score (log 10)')

  abundance_plot <- shared %>% 
  	gather(OTU, abundance, matches('Otu\\d*')) %>% 
  	left_join(taxonomy, by = 'OTU') %>% 
  	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
    left_join(lefse_design, by = c('Group')) %>% 
  	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
  	summarise(abundance = sum(abundance)) %>% 
  	mutate(tax_otu_label = Genus) %>% 
  	mutate(abundance = abundance / 21.07,
  		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
    #highlight_key(~mouse_id) %>% 
  	ggplot(aes(x = reorder(tax_otu_label, order), 
  			y = abundance, color = class)) + # , shape = mouse_id remove mouse_id from interactive plot
  		stat_summary(fun.data = 'median_hilow', aes(group = class),
  			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
  		theme_bw() + labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
  		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
  			labels=c("0","0.1","1","10","100")) +
  		scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
      scale_shape_manual(values = rep(16, 54)) + 
  		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
  		theme(panel.grid.major.y = element_blank(), 
  			panel.grid.minor = element_blank(),
  			legend.position = 'right') + 
  		coord_flip() + 
      guides(shape = 'none')

  cowplot::plot_grid(lefse_plot, 
                     abundance_plot + 
                       geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.25) + 
                       theme(axis.ticks.y = element_blank(),
                             axis.text.y = element_blank()), 
                     nrow = 1)
}

#library(plotly)
#gg <- ggplotly(abundance_plot + geom_jitter(position = position_jitterdodge(dodge.width = 0.5)), tooltip = c('mouse_id'))
#highlight(hide_legend(gg), on = "plotly_click", off = "plotly_doubleclick", 
#           color = "black", selectize = T, opacityDim = .2,
#           selected = attrs_selected(showlegend = F, marker = list(symbol = 'X')))

```
## Severe disease
```{r}
plot_lefse_data('severe_disease')
```

## Toxin presence
```{r}
plot_lefse_data('toxin_presence')
```

## Summary score
```{r}
plot_lefse_data('summary_score_hilo')
```

## Epithelial damage
```{r}
plot_lefse_data('epithelial_damage_presence')
```
